Applying the metallicity-dependent binary fraction to double white dwarf formation: \\ Implications for LISA

### 2021-11-26 src/ms.tex

\section{Introduction} \label{sec:intro}
Most stars in the Galaxy will end their lives as white dwarfs. Of the stars which are born with a binary companion, many will undergo interactions which bring the two stars closer together, eventually forming a close double white dwarf (DWD). Close DWDs, with orbital periods shorter than $\sim2.75\,\rm{hr}$ %\snotes{[close DWDs are defined to have $P_{\rm{orb}}\leq10^4$ days - are we noting 2.75 hrs for GW sources specifically? If so maybe we should reword this bit because it reads like it conflicts with the abstract?]},
%Katie: Close DWDs is a different classification than close binaries.
are the largest source by number of mHz gravitational waves (GWs) in the Galaxy \citep[e.g.,][]{LISAMissionProposal}. The Laser Interferometer Space Antenna (LISA) is expected to resolve at least $10^4$ individual DWD binaries in the Milky Way and will also observe GW emission from the entire Galactic DWD population through the unresolved foreground created by overlapping signals at sub-mHz frequencies \citep[e.g.,][]{Nelemans2001a, Ruiter2010, Nissanke2012, Yu2013, Korol2017, Lamberts2019, Breivik2020a}. The resolved population will enable the study of several important aspects of binary evolution like the strength of tides \citep{Valsecchi2012} and the stability of mass transfer in DWD systems \citep[e.g.,][]{Marsh2004, Shen2015, Gokhale2007, Sepinsky2014, Kremer2015} as well as provide a probe of Galactic structure \citep{Korol2019} and the Local Group \citep{Korol2018}. The shape and strength of the Galactic DWD foreground can also be used as a tool to study the structure of the Milky Way \citep{Benacquista2006, Breivik2020b} as well as the separation distribution of close DWDs \citep{Korol2021}.

@@

%These relatively close binaries are likely to form from fragmentation, accretion, and inward migration in protostellar disks. This is in contrast to wide binaries whose formation channels stem from fragmentation of molecular cores, which is generally unaffected by the opacity of the material. 
%This is depicted in the ``metallicity invariance" for the IMF and binary fraction beyond separations of 200 au for -1.0 $<$ [Fe/H] $<$ 0.5 of Mazzola et al. (2020). Within protostellar disks however, it has been shown through analytical models and hydrodynamic simulations that the probability of disk fragmentation decreases with increasing metallicity (Mazzola et al. 2020). Metal-poor cores are unable to radiate through molecular transitions as effectively, and so are generally hotter and must reach higher masses in order to collapse into disks, which instigates higher accretion rates and results in gravitational instability and thus fragmentation. Furthermore, if you decrease a disk’s metallicity and thus its optical depth, the mid-plane of the disk is able to radiate and cool more efficiently which encourages fragmentation. (Mazzola et al. 2020 and others).

%Numerically, the dominant sources for LISA will be DWDs from the Milky Way. There is an expectation that several tens of millions of detached DWDs will be present in the LISA band with about ten thousand of these to be individually resolvable  with $f_\text{GW} \geq 0.4$ mHz (Nelemans et al. 2001b). DWDs are nice systems to work with: a compact binary which emits continuously and near monochromatically GW signals in the source frame (LISA proposal REF), and of which the second harmonic of their signals can be measured with high accuracy. In general, higher frequency systems are louder and better characterized than low frequency counterparts. In fact, at low frequencies we find that the high number of DWDs begin to overlap in their signals, creating a stochastic background. This effective “confusion signal” arises for these low frequencies from having more than one DWD per frequency bin, $f_\text{bin} = 1 / t_\text{obs}$, where $t_\text{obs}$ is the length of observation time for the LISA mission, which we assume to be four years (LISA proposal REF).

%The science that the LISA mission will contribute to is diverse in scope. Specific to DWDs, the hope is to use the distribution of chirp masses and periods to constrain the impact of the common envelope, which drastically tightens the orbit of the systems (Toonen et al. 2014). A complete sample will also allow for direct comparison with the post-common envelope binaries, which have undergone only one episode of the common envelope (Rebassa-Mansergas et al. 2012). In some cases, the frequency derivative will be measurable and allow us to determine if mass transfer is happening (Breivik et al. 2018) and/or tidal interactions are deforming the white dwarfs.

@@

%The Latte suite of FIRE-2 cosmological zoom-in baryonic simulations of Milky Way-mass galaxies \citep{Wetzel2016}, part of the Feedback In Realistic Environments (FIRE) simulation project, were run using the Gizmo gravity plus hydrodynamics code in meshless finite-mass (MFM) mode \citep{Hopkins2015} and the FIRE-2 physics model \citep{Hopkins2018}. Synthetic Gaia DR2-like surveys of the Latte suite of FIRE-2 simulations were created via the Ananke framework \citep{Sanderson2020}.

\subsection{Binary population models}
\label{sec:bin_pop}
We simulate the evolution of DWD progenitor populations using \cosmic\footnote{https://cosmic-popsynth.github.io}, an open-source Python-based rapid binary population synthesis suite which employs single and binary star evolution using \texttt{SSE/BSE} \citep{Hurley2000, Hurley2002}. Several modifications have been added to \cosmic\ which incorporate updates for massive star evolution and binary interactions. For a detailed description of these modifications see \citet{Breivik2020a}. \cosmic\ has been used in several studies to examine the effects of binary evolution on binary populations from blue stragglers \citep{Leiner2021} and heartbeat stars \citep{Jayasinghe2021}, to white dwarf populations \citep{Kremer2017,Breivik2018,Kilic2021}, to merging compact object populations in isolated binaries \citep{Zevin2020b, Zevin2020a, Zevin2021, Wong2021, Mandhai2021} and in dynamical environments around super-massive black holes \citep{Stephan2019, Wang2021}. 
%Recently, \cosmic\ was integrated with the Cluster Monte Carlo (CMC) code to enable direct comparisons between isolated binary evolution and the evolution of stars in globular clusters \citep{Rodriguez2021}. Since this work considers the effects of the metallicity-dependence of the close isolated binary fraction, we leave direct comparisons to globular clusters to future work.

@@

% I'm suggesting text above since we have a boiler-plate style that we use to stay consisten in the literature.
%Rather than evolving an initial binary population’s individual stars - a computationally expensive process - \cosmic\ uses Monte Carlo statistics to create fixed populations that would be expected based on parameters like SFH and the known processes within binary star evolution. In the \cosmic\ framework, binaries are initialized iteratively with zero age main sequence (ZAMS) binary parameters and tested against the final parameter distributions desired by the user until they match. The output of \cosmic\ contains enough information that a user can analyze both individual systems of interest as well as the overall population, from ZAMS until compact binary formation.

%In this study, we used \cosmic\ 's cosmic-pop package to create fixed populations of DWDs, keeping all parameters except DWD type and metallicity Z constant. The output fully describes the binary parameter distributions that result from our SFH and binary evolution model, coined BSEDict, which is specified in an inifile containing sections on filters, convergence, sampling, and BSE flags, which will be described below:

%\katie{Based on the restructuring I did, I think you should expand this paragraph to describe (1) since we are interested in a metallicity-dependent binary fraction, we simulate metallicity bins separately, (2) since the masses and orbital periods of each DWD type at formation span a wide range, we simulate them separately.}
%For a given metallicity, a fixed population of one million binaries was created for each DWD subtype (discussed in Section \ref{sec:ini}). 
%\katie{Is this true? I think we probably simulated more. We should talk about this today!}

%moved paragraph up here from below since it takes care of some of the things I suggested in the paragraph above!

@@

%this is covered in my text above.
%We iteratively simulate binary populations until the distributions of the DWD masses and orbital period at the formation of the second-formed WD converge to a stable shape. Eccentricity does not need to be analyzed, as most DWDs are circularized by tidal forces (Marsh et al. 2004, Gokhale et al. 2007, Kremer et al. 2015/17), which also simplifies many calculations later on for GW evolution and GW signals.

@@

\subsection{A metallicity-dependent SFH: Convolving with the FIRE models}
\label{sec:FIRE}
%\katie{I think this paragraph gives a bit too much detail on FIRE. I think the only detail we need is (1) we use Galaxy m12i from the Latte Suite of the FIRE simulations and (2) Latte reproduces parts of the Milky Way (see Lamberts+2019 paper for this info)}
To create Milky Way-like galaxies which integrate the metallicity-dependent binary fraction, we use the metallicity-dependent ages and positions of galaxy \textbf{m12i} from the ``Latte" suite of the FIRE simulations \citep{Hopkins2015, Wetzel2016, Hopkins2018} to create synthetic, Milky Way-like DWD populations. %\sout{Specifically, the new “FIRE-2” code version was used, which updates the FIRE physics modules in the code GIZMO (Hopkins et al. 2018;  Gabyurov and Nitadori 2011; Hopkins 2015). This update leverages recent developments of mesh-free Godunov hydrodynamics methods, and incorporates updates to “cooling and recombination rates, gravitational force softening, and numerical feedback coupling” (Hopkins et al. 2018). The Latte suite is then run using Gizmo gravity, plus hydrodynamics code in meshless finite-mass (MFM) mode (Hopkins 2015) and the FIRE-2 physics model (Hopkins e al 2018)” $<$- citation}.

@@

The metallicity-dependent binary fraction is shown in black in Figure \ref{fig:sfh_vs_fb} along with the mass in star particles from galaxy \textbf{m12i}, shown in red, as a function of our metallicity grid. %Plotted in black is the binary fraction as a function of metallicity. 
%The metallicity values chosen for our grid are shown as the scatter points. The red histogram depicts the total mass in stars formed for each metallicity bin within FIRE galaxy \textbf{m12i}, the Milky Way-like simulation used in this study which will be described further in Section \ref{sec:simulations}, as a function of metallicity. 
The binary fraction, $f_{\rm{b}}$, drops drastically across metallicity while the mass formed in \textbf{m12i} increases significantly. These two opposing trends compete throughout this study along with the impact of metallicity on single star evolution to form the final numerical distribution of systems in our DWD populations.

@@

%We use these ratios to then determine the total amount of ZAMS mass sampled in single and binary stars. 
%This ratio can then be used to scale the total amount of ZAMS mass sampled in binary stars required to produce our simulated population from \cosmic\ to $M_{\rm{ZAMS,sim}}$, the total amount of ZAMS mass in single stars and binary systems required to produce the simulated population for a given binary fraction $f_{\rm{b}}$.

%For a given metallicity in our grid and a given DWD type of one compact object binary type, the general procedure to create the Milky Way like galaxy is as follows. We specify metallicity bin $i$ with metallicities Z $\in$ [Z$_{i-1}$, Z$_i$]. The number of DWDs per unit mass N$_{\text{DWD,}i}$ for the \katie{\sout{cosmic-pop} simulated \cosmic} population pop$_i$ with Z = Z$_i$ is found through:

%\begin{align*}
%    R_{m}(f_i) &= \frac{M_{\text{singles,}\odot}(f_i)}{M_\text{binaries,} \odot(f_i)} \numberthis \\
%    \\
%    M_{\text{tot,}i} &=  M_{\text{binaries,}i} \cdot (1 + R_m(f_i)) \numberthis \\
%    \\
%    N_{\text{DWD,}i} &= \frac{N_{\text{DWD, tot,}i}}{M_{\text{tot,}i}} \numberthis
%\end{align*}

%R$_m$ is the fraction of mass in single stars to mass in binary stars for binary fraction $f_i$. \katie{I wonder if we can rewrite these equations to be a bit less confusing; I think we should use Lamberts+2019 as a guide.} \katie{\sout{This is calculated ahead of time using the \cosmic\ independent sampler in the InitialBinaryTable class.} I think that this sentence should say that the scaling incorporates the same assumptions for primary and secondary mass distributions as are used in the simulated data.} One million DWD binaries are initialized at solar metallicity for a given binary fraction, along with the total mass in singles vs. binaries. M$_{{\rm{binaries, }}i}$ the mass in binaries of pop$_i$ (the total mass of the population, as they were run with f$_{\rm{b}}$ = 1.0). M$_{{\rm{binaries, }}i}$ is scaled using R$_m$ to find the total astronomical mass M$_{\rm{tot, Z}_i}$ of the population. N$_{{\rm{DWD, tot,}}i}$ is the total number of DWDs in pop$_i$.

@@

%Already covered above
%In order to ensure that each DWD position is unique, we assign positions inside the star particle following the Ananke framework which provides offsets from the star particle center.

%\katie{I think this should go later}This gives us a final distance $d$ to the sun for all the systems, and from this and other information, we can calculate parameters relevant to the LISA observatory like Amplitude Spectral Density (ASD), signal-to-noise ratio (SNR), chirps, chirp masses, etc, as discussed in the next section.

%This will be a floating point number. The integer component gives one number of DWDs to sample with replacement from the cosmic-pop output to apply to each FIRE star particle. The decimal component will be treated as a probability $p$ of a binary existing for that particle. We assign an additional DWD to each star particle with probability $p$. The DWDs now have metallicities, ages, and galactic positions.

%\katie{I think the presentation of the next couple of paragraphs is a bit out of order. What about presenting it in the order: (1) we evolve the DWDs from their formation time up to the present day using the Peters equations (2) we discard any systems which overflow their roche lobes.}

@@

%Following the creation of an initial population, certain systems are discarded. We discard any systems with formation times longer than their assigned age from their respective star particle since these systems won’t have formed a DWD by present day and would thus not be observable by LISA. The time from DWD formation until the time when the system merges, t$_{\text{merge}}$, is defined as 
%\begin{equation}
%    t_\text{merge} = \frac{a_i^4}{4 \beta}
%\end{equation}
%where $a_i$ is the DWD separation at formation, and 
%\begin{equation}
%    \beta = \frac{64G^3}{5c^5} M_1M_2(M_1+M_2)
%\end{equation}
%is constant throughout DWD evolution \citep{Peters1964}. The delay time of the %system, $t_\text{delay}$, is defined as the sum of the DWD formation time and t$_{\text{merge}}$. Systems are also discarded if the delay time is less than the FIRE age, since they would have merged before present day. Finally, we 
%We find the time corresponding to when} the system will reach separation $a = a_\text{RLOF,2}$ by inverting Eq.(12), and \textcolor{purple}{\sout{discarding systems with $t_\text{phys} + t_\text{RLOF} \leq t_\text{FIRE}$} discard systems that would overflow their Roche Lobe before present day according to their FIRE age}.

%For the non-discarded systems, we assume evolution occurs solely through the emission of gravitational radiation. Furthermore,

%We discard any sampled DWDs for which the sum of their ZAMS birth time given by the star particle and DWD formation time is larger than the age of the star particle. 
%since DWDs orbit mainly in the second harmonic, with negligible signal stemming from the others. Systems in the LISA band are defined to have $f_\text{GW} \in [10^{-4}, 1.0]$ Hz, although they may not be observable from Earth with LISA’s current sensitivity limits.

%\katie{I think we might want to leave photometric observations to a future paper}
%In order to investigate possible follow-up with photometric observations, we also calculate the bolometric magnitude of the DWD. To calculate the luminosity of the DWD, we utilize the modified Mestel cooling described in Hurley(2003):
%\begin{equation}
%    L_{\text{WD,}i} = %\frac{bM_iZ^{0.4}}{(A_i(t_\text{evol,i}+0.1))^x}
%\end{equation}
%where $M_i$ is the WD mass in solar masses and $Z$ is the metallicity of our specified metallicity bin. $A_i$ is the baryon number for the WD subtype. 4 corresponds to helium, 15 to carbon-oxygen and 17 to oxygen-neon. $t_\text{evol, i}$ is evolution time for WD component $i$ in Myr (i.e. $t_{\text{evol, 2}}$ = $t_{\text{evol}}$, but $t_{\text{evol,1}} = t_{\text{FIRE}}-t_{\text{phys,1}}$ where $t_{\text{phys,1}}$ must be retrieved from \cosmic\ at primary formation.). The values of $b$ and $x$ are dependent on this age:
%\begin{equation}
%x = \begin{cases} 
%     1.18 & t_\text{evol, i} < 9000.0\\
%      6.48 & t_\text{evol, i} \geq 9000.0
%   \end{cases}
%\end{equation}
%\begin{equation}
%b = \begin{cases} 
%      300 & t_\text{evol, i} < 9000.0\\
%      300(9000A)^{5/3} & t_\text{evol, i} \geq 9000.0 
%   \end{cases}
%\end{equation}
%We can then find the bolometric magnitude of the system using https://www.astro.princeton.edu/~gk/A403/
%constants.pdf and https://www.astro.keele.ac.uk/jkt/pubs/
%JKTeq-fluxsum.pdf:
%\begin{equation}
%    m_{\text{bol, tot}} = -2.5 \cdot \log_{10}(10^{-0.4m_1}+10^{-0.4m_2}),
%\end{equation}
%where the bolometric magnitude of each component $m_{i}$ is %given by:
%\begin{equation}
%    m_{i} = 4.8 - 2.5 \cdot %\log_{10}\left(\frac{L_i}{L_\odot}\right)
%\end{equation}

@@

\noindent where $h_c$ is the characteristic strain of the system, $S_n(f)$ is the LISA sensitivity curve of \citet{Robson2019}, and the frequency limits are determined by the orbital evolution over the observation time, $T_{\rm{obs}}=4\,\rm{yr}$. The characteristic strain for %DWDs 
circular orbits is 
%only defined for the $n=2$ harmonic and is:
\begin{equation}
    h_{c}^2 = \frac{2^{2/3}}{3 \pi^{4/3}} \frac{(G \mathcal{M}_c)^{5/3}}{c^3 D_L^2} \frac{1}{f_{\rm orb}^{1/3}},
\end{equation}
\noindent where $\mathcal{M}_c = (M_1M_2)^{3/5}/(M_1 + M_2)^{1/5}$ is the system's chirp mass, and $D_L$ is the systems luminosity distance which we assume to be the distance of each simulated DWD to the Sun.

@@

%$g(n,e)$ is the relative power of gravitational radiation at the nth harmonic taken from Peters and Mathews(1963). For $n=2$ this reduces to:
%\begin{equation}
%    h_2^2 = \frac{2^{22/3}}{5}\frac{(G\mathcal{M}_c)^{10/3}}{c^8D_L^2}(\pi f_\text{orb})^{4/3}
%\end{equation}
The amplitude spectral density for a stationary system is finally defined as $ASD = h_{2} \sqrt{T_\text{obs}}$, such that the SNR for stationary source is simply, $\rho \sim ASD/S_n$.

@@

\subsection{DWD types and their formation channels}\label{sec:ini}
As discussed in Section\,\ref{sec:bin_pop}, we consider four DWD sub-types, which each contribute differently to LISA's GW signals: He + He, CO + He, CO + CO, and ONe + X. Each sub-type has a unique distribution in their formation times, initial masses, radii, and orbital periods stemming from variations in their evolution channels and their formation efficiency. Here we describe the general formation scenarios and population properties of Galactic close DWDs which may be observable by LISA.
%We thus simulated \cosmic \ DWD populations separately for each DWD type and metallicity value in order to identify variations in how each type is impacted differently by changes in metallicity. A key effect is that in general, lower metallicity stars have shorter MS lifetimes and smaller initial radii than high metallicity counterparts (https://arxiv.org/pdf/2001.10404.pdf). This is discussed in greater depth later in this section.

%The resulting DWD populations from \cosmic\ gives insight in to the impact that metallicity has on these systems upon their initial formation. The trends we see within these distributions can be better interpreted through the characteristics and evolutionary channels of each DWD type which are outlined below.

%He + He: (REF = https://iopscience.iop.org/article/
%10.1086/303686/fulltext/34537.text.html)
He WDs are unable to form through single star evolution within the lifetime of the Milky Way. Instead, they originate through interactions in close binary systems or binaries with large eccentricities. Because of this, He WDs are able to form with low component masses on order $\sim$ 0.1M$_\odot$, with the majority of He WDs in our simulations having masses between 0.2-0.5 M$_\odot$. 
%, which would be impossible through SSE within a Hubble time (REF). 
He + He DWDs form through evolution of close binary systems, during which their envelopes are both stripped through RLO and CE phase interactions before Helium ignition occurs. The two progenitor stars generally have masses $\lesssim 3 M_\odot$ which is lower than the progenitors of other DWD types. Our simulated He + He DWDs have an approximately constant distribution of formation times $\gtrsim$ 2.5 Gyr. Lastly, since the ZAMS separations are skewed towards shorter values, we also see that the resulting DWD separations are smaller on average than that of other DWD types.

@@

%CO + He: 
%(REF = https://www.researchgate.net/publ ication/231115182\_Helium\_and\_Carbon-Oxygen\_White\_ Dwarfs\_in\_Close\_Binaries)
A CO WD forms when a star is able to begin the helium burning process before its envelope is stripped. Thus to form a CO + He DWD, RLO and CE stages occur after one component experiences core helium burning, but before the other component can. Most close CO + He DWDs form in approximately 2 Gyr after ZAMS and with very short periods because the He WD is formed through the ejection of a common envelope which greatly reduces the orbital separation. Because of these short formation separations, many CO + He DWDs merge before the present day.
%(Lamberts) of $O$(hour), and may merge due to GW radiation by present day. Specifically, DWDs with sub-hour periods tend to merge in less than 10$^8$ years (REF). 
Due to their asymmetric mass distributions, they have lower chirp masses, but their shorter periods make them important candidates for LISA detection. 
%Possible formation channels of these short period CO + He DWDs are explored in (REF). The fate of the binary is dependent on the orbital period that results from the final CE phase before DWD formation.

%CO + CO:
%A CO WD forms when a star is able to begin helium burning before its envelope is stripped (Lamberts,  https://arxiv.org/pdf/2011.10439.pdf). 
To prevent the two stars' envelopes from being stripped before helium ignition, CO + CO DWDs typically form from progenitors in wider orbits, and the two components may have little to no interaction during their evolution from ZAMS to DWD. CO DWDs thus have a distribution in progenitor separation that extends to larger values than for other DWD types. Most need $> 0.3$ Gyr to form. These systems have component WD masses between $0.35$--$1.0\,M_\odot$ and make up the majority of the DWD population.

%ONe + X: (REF = https://arxiv.org/pdf/2011.10439.pdf)
ONe WDs are rare and typically form from massive progenitor stars which evolve through the asymptotic giant branch (AGB) phase, thus resulting in a higher mass WD. All ONe WDs in our \cosmic\ populations, e.g., have progenitor ZAMS masses above $4\,M_\odot$, and the resulting ONe WDs have a relatively flat distribution of masses from $1.05\,M_\odot$ up to the Chandrasekkhar limit of $1.4\,M_\odot$. Because an ONe WD can have a companion of any other WD type in our study, there is a spread in their distributions for separation, secondary mass, final orbital period, and formation time. In general, however, these systems result from wider separations to allow for the evolution of the ONe component without merging, e.g. all initial separations in our \cosmic \ populations have separations $\gtrsim 1.5\,R_\odot$. ONe + X DWDs can form on short timescales, as low as $30\,\rm{Myr}$ for the majority of high-metallicity systems.

@@

%KB cut 28 sept 21
%This drop across metallicity is due to an increase in the number of merging systems before the formation of the DWD.
%The increased propensity for merging during binary interactions stems from variations in the star's intrinsic properties caused by the effect of metallicity on stellar evolution. A higher abundance of metals results in a cooler, dimmer star with a lower central temperature and that is less compact, leading to larger stellar radii and longer main sequence lifetimes. These changes further alter the timescales for Roche-lobe overflow interactions.

%Consider two stars of high and low metallicity but equal mass. The star with higher metal abundance will have a larger radius, and a higher fraction of mass in the convective envelope (which creates longer turnover timescales) of the star/less in the core. This in theory impacts the CE phase - there would be more envelope to eject, leading to a longer energy transfer times, longer CE phases, and thus more mergers. We will now detail the unique processes for each DWD type which leads to the drop in DWD formation efficiency. For many of the subtypes, the mechanisms behind and degree to which the DWD formation efficiency is impacted can be identified within subsets of their initial orbital period.
%\katie{I think we might want to shorten this previous paragraph a bit since we go into detail later on.}

For He + He DWDs, the sudden drop in formation efficiency near $\log(Z/Z_{\odot})=-1.0$, is generally caused by the timescale for which the initially more massive star in the DWD progenitor overflows it's Roche lobe. At lower metallicities, donors tend to fill their Roche lobes while they are still on the main sequence and the mass transfer remains stable. This is because, in our \cosmic\ models %\snotes{[in just our \cosmic\ models or astrophysically? or are we stating this is just for \cosmic\ to be safe?]}, 
% KB: yeah exactly -- to be safe
lower metallicity stars evolve faster than high-metallicity stars and have larger radii near the end of the MS. At higher metallicities, mass transfer is initiated when the donor has left the main sequence and the binary enters a common envelope evolution. The ZAMS orbital period leads to different specific evolutionary channels. For short-period systems with periods below $10~\rm{days}$, this leads to a stellar merger due to insufficient orbital energy to eject the envelope. Stellar mergers continue to dominate the evolutionary pathways of systems with intermediate orbital periods ($1 < \log_{10}(P_{\rm{orb}}/\rm{day}) < 2.5$). However, an additional growing number of merging systems arises with one WD component and one stellar companion, and systems which don't interact at all and thus do not form a He + He DWD. The distinction between the various scenarios in this intermediate orbital period range depends on the combination of their ZAMS masses and orbital periods. Finally, at wider initial periods $\log_{10}(\rm{P}) > 2.5$, the decrease in formation efficiency is dominated by systems which never interact and thus do not form a DWD before the present day.

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/CEsep.pdf}
    \caption{Average interaction separation, $\overline{\mathrm{a}}_{\rm{CE}}$ of progenitors of close DWDs across metallicity for each DWD type from model FZ. Solid lines show the average separation at the first RLO for binaries in each metallicity bin. The shaded regions show the $1\sigma$ spread around the mean within each metallicity bin. The average interaction separation increases with metallicity for every DWD type. The positive trend in the average interaction separation is a direct consequence of larger envelope masses of higher-metallicity donors which are less evolved than their lower-metallicity counterparts. The sharp drop of the CO + CO average at the highest-metallicity bin is due to binning effects of the figure: it only contains 115 systems, of which the majority stem from a single low-separation system which was sampled 94 times times during the simulation.} 
    %The initial radii of high-metallicity ZAMS stars is larger than their lower-metallicity counterparts but inflate less during their evolution, creating three cases: 1) stars with wider initial separations that do not shrink enough to evolve into the LISA band before present day or fail to interact at all since their radii never increase enough to interact, 2) systems which remain stable throughout their evolution and end up in the LISA band, which requires higher interaction separations to account for the larger radii and deeper convective envelopes which need to be ejected (i.e the separation must be large enough to allow for longer CE phases), and 3) systems which merge before present day either through interactions delayed until late in their evolution from radii which do not evolve much, causing unstable mass transfer, or through very close initial separations which merged during a CE phase before present. Only case (2) would result in successful systems shown here.}
    \label{fig:CEsep}
\end{figure*}

>>>>

\section{Introduction} \label{sec:intro}
Most stars in the Galaxy will end their lives as white dwarfs. Of the stars which are born with a binary companion, many will undergo interactions which bring the two stars closer together, eventually forming a close double white dwarf (DWD). Close DWDs, with orbital periods shorter than $\sim2.75\,\rm{hr}$ are the largest source by number of mHz gravitational waves (GWs) in the Galaxy \citep[e.g.,][]{LISAMissionProposal}. The Laser Interferometer Space Antenna (LISA) is expected to resolve at least $10^4$ individual DWD binaries in the Milky Way and will also observe GW emission from the entire Galactic DWD population through the unresolved foreground created by overlapping signals at sub-mHz frequencies \citep[e.g.,][]{Nelemans2001a, Ruiter2010, Nissanke2012, Yu2013, Korol2017, Lamberts2019, Breivik2020a}. The resolved population will enable the study of several important aspects of binary evolution like the strength of tides \citep{Valsecchi2012} and the stability of mass transfer in DWD systems \citep[e.g.,][]{Marsh2004, Shen2015, Gokhale2007, Sepinsky2014, Kremer2015} as well as provide a probe of Galactic structure \citep{Korol2019} and the Local Group \citep{Korol2018}. The shape and strength of the Galactic DWD foreground can also be used as a tool to study the structure of the Milky Way \citep{Benacquista2006, Breivik2020b} as well as the separation distribution of close DWDs \citep{Korol2021}.

@@

@@

\subsection{Binary population models}
\label{sec:bin_pop}
We simulate the evolution of DWD progenitor populations using \cosmic\footnote{https://cosmic-popsynth.github.io}, an open-source Python-based rapid binary population synthesis suite which employs single and binary star evolution using \texttt{SSE/BSE} \citep{Hurley2000, Hurley2002}. Several modifications have been added to \cosmic\ which incorporate updates for massive star evolution and binary interactions. For a detailed description of these modifications see \citet{Breivik2020a}. \cosmic\ has been used in several studies to examine the effects of binary evolution on binary populations from blue stragglers \citep{Leiner2021} and heartbeat stars \citep{Jayasinghe2021}, to white dwarf populations \citep{Kremer2017,Breivik2018,Kilic2021}, to merging compact object populations in isolated binaries \citep{Zevin2020b, Zevin2020a, Zevin2021, Wong2021, Mandhai2021} and in dynamical environments around super-massive black holes \citep{Stephan2019, Wang2021}.

@@

@@

@@

\subsection{A metallicity-dependent SFH: Convolving with the FIRE models}
\label{sec:FIRE}
To create Milky Way-like galaxies which integrate the metallicity-dependent binary fraction, we use the metallicity-dependent ages and positions of galaxy \textbf{m12i} from the ``Latte" suite of the FIRE simulations \citep{Hopkins2015, Wetzel2016, Hopkins2018} to create synthetic, Milky Way-like DWD populations.

@@

The metallicity-dependent binary fraction is shown in black in Figure \ref{fig:sfh_vs_fb} along with the mass in star particles from galaxy \textbf{m12i}, shown in red, as a function of our metallicity grid. The binary fraction, $f_{\rm{b}}$, drops drastically across metallicity while the mass formed in \textbf{m12i} increases significantly. These two opposing trends compete throughout this study along with the impact of metallicity on single star evolution to form the final numerical distribution of systems in our DWD populations.

@@

@@

@@

@@

\noindent where $h_c$ is the characteristic strain of the system, $S_n(f)$ is the LISA sensitivity curve of \citet{Robson2019}, and the frequency limits are determined by the orbital evolution over the observation time, $T_{\rm{obs}}=4\,\rm{yr}$. The characteristic strain for %DWDs 
circular orbits is 
\begin{equation}
    h_{c}^2 = \frac{2^{2/3}}{3 \pi^{4/3}} \frac{(G \mathcal{M}_c)^{5/3}}{c^3 D_L^2} \frac{1}{f_{\rm orb}^{1/3}},
\end{equation}
\noindent where $\mathcal{M}_c = (M_1M_2)^{3/5}/(M_1 + M_2)^{1/5}$ is the system's chirp mass, and $D_L$ is the systems luminosity distance which we assume to be the distance of each simulated DWD to the Sun.

@@

The amplitude spectral density for a stationary system is finally defined as $ASD = h_{2} \sqrt{T_\text{obs}}$, such that the SNR for stationary source is simply, $\rho \sim ASD/S_n$.

@@

\subsection{DWD types and their formation channels}\label{sec:ini}
As discussed in Section\,\ref{sec:bin_pop}, we consider four DWD sub-types, which each contribute differently to LISA's GW signals: He + He, CO + He, CO + CO, and ONe + X. Each sub-type has a unique distribution in their formation times, initial masses, radii, and orbital periods stemming from variations in their evolution channels and their formation efficiency. Here we describe the general formation scenarios and population properties of Galactic close DWDs which may be observable by LISA.

He WDs are unable to form through single star evolution within the lifetime of the Milky Way. Instead, they originate through interactions in close binary systems or binaries with large eccentricities. Because of this, He WDs are able to form with low component masses on order $\sim$ 0.1M$_\odot$, with the majority of He WDs in our simulations having masses between 0.2-0.5 M$_\odot$. He + He DWDs form through evolution of close binary systems, during which their envelopes are both stripped through RLO and CE phase interactions before Helium ignition occurs. The two progenitor stars generally have masses $\lesssim 3 M_\odot$ which is lower than the progenitors of other DWD types. Our simulated He + He DWDs have an approximately constant distribution of formation times $\gtrsim$ 2.5 Gyr. Lastly, since the ZAMS separations are skewed towards shorter values, we also see that the resulting DWD separations are smaller on average than that of other DWD types.

@@

A CO WD forms when a star is able to begin the helium burning process before its envelope is stripped. Thus to form a CO + He DWD, RLO and CE stages occur after one component experiences core helium burning, but before the other component can. Most close CO + He DWDs form in approximately 2 Gyr after ZAMS and with very short periods because the He WD is formed through the ejection of a common envelope which greatly reduces the orbital separation. Because of these short formation separations, many CO + He DWDs merge before the present day. Due to their asymmetric mass distributions, they have lower chirp masses, but their shorter periods make them important candidates for LISA detection.

To prevent the two stars' envelopes from being stripped before helium ignition, CO + CO DWDs typically form from progenitors in wider orbits, and the two components may have little to no interaction during their evolution from ZAMS to DWD. CO DWDs thus have a distribution in progenitor separation that extends to larger values than for other DWD types. Most need $> 0.3$ Gyr to form. These systems have component WD masses between $0.35$--$1.0\,M_\odot$ and make up the majority of the DWD population.

ONe WDs are rare and typically form from massive progenitor stars which evolve through the asymptotic giant branch (AGB) phase, thus resulting in a higher mass WD. All ONe WDs in our \cosmic\ populations, e.g., have progenitor ZAMS masses above $4\,M_\odot$, and the resulting ONe WDs have a relatively flat distribution of masses from $1.05\,M_\odot$ up to the Chandrasekkhar limit of $1.4\,M_\odot$. Because an ONe WD can have a companion of any other WD type in our study, there is a spread in their distributions for separation, secondary mass, final orbital period, and formation time. In general, however, these systems result from wider separations to allow for the evolution of the ONe component without merging, e.g. all initial separations in our \cosmic \ populations have separations $\gtrsim 1.5\,R_\odot$. ONe + X DWDs can form on short timescales, as low as $30\,\rm{Myr}$ for the majority of high-metallicity systems.

@@

For He + He DWDs, the sudden drop in formation efficiency near $\log(Z/Z_{\odot})=-1.0$, is generally caused by the timescale for which the initially more massive star in the DWD progenitor overflows it's Roche lobe. At lower metallicities, donors tend to fill their Roche lobes while they are still on the main sequence and the mass transfer remains stable. This is because, in our \cosmic\ models lower metallicity stars evolve faster than high-metallicity stars and have larger radii near the end of the MS. At higher metallicities, mass transfer is initiated when the donor has left the main sequence and the binary enters a common envelope evolution. The ZAMS orbital period leads to different specific evolutionary channels. For short-period systems with periods below $10~\rm{days}$, this leads to a stellar merger due to insufficient orbital energy to eject the envelope. Stellar mergers continue to dominate the evolutionary pathways of systems with intermediate orbital periods ($1 < \log_{10}(P_{\rm{orb}}/\rm{day}) < 2.5$). However, an additional growing number of merging systems arises with one WD component and one stellar companion, and systems which don't interact at all and thus do not form a He + He DWD. The distinction between the various scenarios in this intermediate orbital period range depends on the combination of their ZAMS masses and orbital periods. Finally, at wider initial periods $\log_{10}(\rm{P}) > 2.5$, the decrease in formation efficiency is dominated by systems which never interact and thus do not form a DWD before the present day.

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/CEsep.pdf}
    \caption{Average interaction separation, $\overline{\mathrm{a}}_{\rm{CE}}$ of progenitors of close DWDs across metallicity for each DWD type from model FZ. Solid lines show the average separation at the first RLO for binaries in each metallicity bin. The shaded regions show the $1\sigma$ spread around the mean within each metallicity bin. The average interaction separation increases with metallicity for every DWD type. The positive trend in the average interaction separation is a direct consequence of larger envelope masses of higher-metallicity donors which are less evolved than their lower-metallicity counterparts. The sharp drop of the CO + CO average at the highest-metallicity bin is due to binning effects of the figure: it only contains 115 systems, of which the majority stem from a single low-separation system which was sampled 94 times times during the simulation.} 
    \label{fig:CEsep}
\end{figure*}
